### 2018-07-04 Stickleback_Paper.tex
latest

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{comment}
\usepackage[backgroundcolor=pink,linecolor=red]{todonotes}

@@

%wider audience by more general questions about ... introduction maybe too fishy???

%a question about the genetic basis for adaptstion, how is it maintianed?
% theres alot of debate in the field

% more specifically - sticklback fish

% we want to examine this, This is why our approach is unique
% This is why the field could benefit from this.

%methods - we did this this and this specifically
%results - we found this and this specifically

% discussion get more broad again.

@@

In 2009, Dolph Schluter and Gina L. Conte suggested the "transportation"-hypothesis. 
This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
It suggests freshwater haplotypes are distributed though marine individuals and are continually selected upon in introduced freshwater populations.
The continued selection on freshwater favored alleles,
and introgression between the two sub-species,
allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection rebuilds the freshwater haplotype. \cite{Schluter2009} 
The alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, and their haplotype is passed down directly. 
This hypothesis has been shown to be unlikely due to 
finding a high frequency of freshwater alleles in the ocean,
and almost no freshwater individuals.

Assuming they were correct about transportation, we might ask; 
how much migration and gene flow is to be expected for this relationship to be maintained?
(Other questions)
how does selection, at the genomic level, alter given a range of migration and recombination rates?
Here, we use forward moving simulations to model marine and freshwater populations of stickleback.
We then observe the effect (focusing on migration and recombination, for now) 
of different parameters on local adaptation of introduced freshwater environments. 
We compare to real data and make predictions about realistic parameters.

Simulations are a powerful tool when there is some map between the parameters we want to learn about
and the genetic data we observe in nature. 
We don't know what that map is, so we invert the relationship. 
This means going from genomic data that we see to the parameter values that must be driving the system, 
assuming the model is correct.

@@

\subsubsection{Geography}

@@

\begin{figure}[htb]
	\begin{center}
  		\includegraphics[width=0.8\linewidth]{GeographyDiagram}
  		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 
		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 10.0]. Each 
		lake$_{i}$ is a discrete population connected by migration only through the marine at position $i - 0.5$. The introduced lakes 
		have the same location and selective pressure as the original lakes, but arise at 40K generations, 
		initialized as a copy of all marine individuals.}
  		\label{fig:Geo}
	\end{center}
\end{figure}

\subsubsection{Selective Pressure}

@@

\subsubsection{Genomic Architecture}

In SLiM, the genome is conceptually a linear array of loci defining locations for 
mutations that arise and breakpoints for recombination. 
Genomic elements define a range depicting the types of mutations that can arise, the ratio at which they can arise. 
Although still uncertain about how much of the genome is directly associated with the distinguished phenotypes, 
GWAS has indicated clusters of loci (linkage groups? Operon?) along the genome to be causative. 
To mimic this and compare dynamics of neutral vs. effect under selection, 
we create regions of effect. 
In these genomic elements of effect, we uniformly allow additive, recessive, and dominant, mutation that can push phenotype in either the positive or negative direction to arise. 
Every genomic position outside of these effect regions only allows for selectively neutral mutations to arise.

@@

%This allows us to observe differences in selection at the genomic level.

@@

\begin{figure}[htb]
	\begin{center}
  		\includegraphics[width=01.0\linewidth]{matplotlib/Fst.png}
  		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
		On the left, $F_{st}$ was calculated only on neutral mutation frequencies. On the right, $F_{st}$ was calculated
		only on effect (Impact on phenotype) mutation frequencies. 
		Effect mutations are acted upon by selection as they have an affect on fitness where neutral mutations  }
  		\label{fig:Fst}
	\end{center}
\end{figure}

@@

%FIGURES THAT SHOW THIS:
%-Time to adaptation
\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=01.0\linewidth]{matplotlib/TimeToAdaptation.png}
  		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
		Each point represents a simulation run at some value of $M$. 
		The yellow dashed line is the average of all points at each respective parameter value}
  		\label{fig:TimeToAdaptation}
	\end{center}
\end{figure}

@@

%FIGURES THAT SHOW THIS:
%-MPAA / IND
%-Total/Avg shared
\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.8\linewidth]{matplotlib/MPFAI.png}
  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) throughout the simulation run, for each subpopulation.
		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
		freshwater adapted alleles.
		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
		while the percentage tells us the variance of the haplotype (note: I also have boxplots telling us total num FAA for the average individual that are interesting to look at "matplotlib/MFAI.png")}
  		\label{fig:MPFAI}
	\end{center}
\end{figure}

@@

%FIGURES THAT SHOW THIS:
%- Phenotype Distribution
\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.8\linewidth]{matplotlib/MeanPhenotype1.png}
  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
		All individuals at generation 0 (beginning of the simulation) 
		}
  		\label{fig:MeanPhenotype}
	\end{center}
\end{figure}

@@

\subsubsection*{Low Recombination Causes Clustering}

\begin{comment}
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
%SHOULD PROBABLY JUST DELETE ALL THIS. V UNORGANIZED.
Across migration rates we observe a large difference in:
gene flow, 
$F_{st}$ values,
amount of time until adaptation,
sharing of solutions (haplotytes?),
$SGV$,
and, migration load. 
we run multiple simulations across a range of migration rates starting at $5e-5$ and increasing by
one order of magnitude up to $.05$. 
With the marine and freshwater both at a population size 2000,
this represents exactly $1/10_{th}$, $1$, $10$, and $100$ migrants per generation, respectively.

At low migration rates of $5e-5$ (1/10 migrants per generation), we observed little sharing of genetic solutions and 
almost no freshwater alleles existing in the marine population.
With an MPFAI of $0.0999$ in the original lakes;
each individual in the original lakes had almost exactly $1/10_{th}$ of the total FAA. 
In addition to each allele appearing in an average of $1.08$ lakes, 
this tells us that each of the 10 lakes has individually selected their own set of solutions (haplotypes) 
rather than sharing alleles through the marine. 
This is further supported by the low capacity of the marine to transport the alleles with an MPFAI of $0.003$.
$T_{adapt}$ remained on an ecological timescale of $18K$ generations, and the correlation of FAA frequencies was at $0.09$.
With the introduced population having an MPFAI of $0.02$, 
we see little evidence that the lakes were able the select on the alleles found in that marine. 
it may be important to note that while we are sampling the 10 introduced lakes as a whole population,
It is possible that only a couple of the lakes even received a copy some alleles during the introduction.
However, to see evidence of the haplotypes being carried throughout the marine we would expect to observe a pattern off all lakes
being able to select upon similar alleles. 
This observation gives a migration rate below the threshold needed to support the transportation hypothesis.

scaling up migration rates to $0.0005$ (1 migrants per generation)
we observe the first evidence of transportation.
The mean total number of FAA drops to 45 and the average number of lakes which each alleles appears in jumped to $1.33$. 
This shows us the gene flow between populations allowed for some freshwater alleles to travel through the marine and be shared by the lakes. 
Interestingly, with the marine showing an MPFAI of $0.02$, 
the introduced was much more able to select upon the alleles;
The introduced population showed a MPFAI of $0.078$, compared to the original lakes which was at $0.129$,
Looking at the FAA frequencies in the two freshwater populations at time of adaptation, we see a correlation of $.68$
While this is evidence of parallel local adaptation, the  $T_{adapt}$ was still at $8,365$ generations.

At $0.005$ (10 migrants per generation) we experience high correlation between the FAA frequencies
in the original and introduced populations, $0.96$, and rapid local adaptation of the introduced population, (81 generations). 
With the introduced population showing an MPFAI of  ~ $0.34$ which is strikingly close to the original freshwater population which is at $0.34$
This migration rate clearly shows the capacity to share alleles in order for us to see rapid and parallel adaptation.

Interestingly, is that if we look at the distribution and total amount of FAA in the marine population at time of introduction between 
$0.0005$ and $0.005$, they are virtually identical.
This comes with two questions
(1) why is there a similar amount of FAA in the marine with such a substantial difference in migration, and
(2) what is it about the higher migration rate of  $0.005$ that allows it perform so much better in terms of rapid adaptation? 
To answer the first question we can see that the median number of defined FAA is much higher, 45
While with the higher migration there is only a median of 10. 
This is a result of the solutions not being shared across the original lakes at lower migration rates.
This can be seen by the average number of lakes each FAA appears in being 
$1.54$ at 1 migrant per generation and 
$3.75$ at 10 migrants per generation.
consequently more alleles were defined as FAA and sampled as such in the marine population. 
This however does not explain why, at higher migration rates, 
the introduced freshwater population was able to rebuild the haplotype so much more quickly. 
For this, It is important to note that after introduction, there is continued migration. 
This means that for a migrate of $0.005$ there is roughly 10 times as 
much of the FAA being brought into the freshwater every generation compared to migrate $0.0005$.

This seams to suggest that continued migration is quite important 
for rapid adaptation at this migration rate as opposed to the simply the concentration of 
freshwater adapted alleles found in the marine.

Note: we should probably investigate the distribution of effect size more to clearly answer the above questions ...

(NOTE:PETER/BILL how does this compare to what we see in nature

\subsection{Qualitative Threshold}

\subsection{Migration Load}

%\begin{figure}
%	\begin{center}
%  		\includegraphics[width=1.2\linewidth]{MigrationLoad}
%  		\caption{A boat.}
%  		\label{fig:boat1}
%	\end{center}
%\end{figure}

While both of the higher migration rates of $0.005$ and $0.05$ showed evidence of sharing alleles and rapid adaptation,
we see substantial difference in the median phenotype compared to the optimum for each subpopulation. 
This is because migration tends to oppose the natural divergence of populations undergoing ecological speciation. \cite{Bolnick}
Migration load depreciation of fitness between two subpopulations with a high amount of introgression between them. 
This is often observed in nature when the sharing of genes constricts the adaptation of both species.

In our simulations, at high migration rates of $0.05$, 
we consistently saw this effect in all subpopulations.

\subsection{Low Recombination in Effect Regions}

Goldilocks Zone?

at 10 migrants/generation (or 0.005 of the population) 
is the first parameter value at which we observe

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\end{comment}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{comment}
\usepackage[backgroundcolor=pink,linecolor=red]{todonotes}
\usepackage{fullpage}

@@

@@

In 2009, Dolph Schluter and Gina L. Conte suggested the "transporter"-hypothesis. 
This outlined the flow of freshwater alleles into marine populations through offspring of hybridization events.
It suggests freshwater haplotypes are distributed though marine individuals and are continually selected upon in introduced freshwater populations.
The continued selection on freshwater favored alleles,
and introgression between the two sub-species,
allows the marine to maintain the freshwater haplotype dispersed in low frequency among its' individuals. 
Once a new freshwater environment is introduced and inhabited by marine individuals who carry freshwater adapted alleles, selection rebuilds the freshwater haplotype. \cite{Schluter2009} 
The alternative to this hypothesis is that individuals from other freshwater environments migrate directly to the new environments, and their haplotype is passed down directly. 
This hypothesis has been shown to be unlikely due to 
finding a high frequency of freshwater alleles in the ocean,
and almost no freshwater individuals.

Assuming they were correct about transportation, we might ask; 
how much migration and gene flow is to be expected for this relationship to be maintained?
%(Other questions)
%how does selection, at the genomic level, alter given a range of migration and recombination rates?
Here, we use forward moving simulations to model marine and freshwater populations of stickleback.
We then observe the effect (focusing on migration and recombination, for now) 
of different parameters on local adaptation of introduced freshwater environments. 
We compare to real data and make predictions about realistic parameters.

@@

\subsubsection*{Geography}

@@

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{GeographyDiagram}
  		\caption{A representation of the geographic and evolutionary history of all populations throughout the simulation. 
		The marine is a one-dimensional, continuous population with spatial positions ranging from [0.0 - 10.0]. Each 
		lake$_{i}$ is a discrete population connected by migration only through the marine at position $i - 0.5$. The introduced lakes 
		have the same location and selective pressure as the original lakes, but arise at 40K generations.
		the introduced population is initialized as a copy of all marine individuals to model marine 
		individuals inhabiting a newly created freshwater environment such as the ponds around Middleton Island.
		We then observe the selection process of marine individuals and following generations 
		with the new selective pressure across a range of parameter values. 
		%Migration between the marine and original lakes is 
		%set to $0$ the generation before introduction to avoid direct migrants from original $->$ introduced lakes.
		}
  		\label{fig:Geo}
	\end{center}
\end{figure}

\subsubsection*{Selective Pressure}

@@

\subsubsection*{Genomic Architecture}

In SLiM, the genome is conceptually a linear array of loci defining locations for 
mutations that arise and breakpoints for recombination. 
Genomic elements define a range depicting the types of mutations that can arise, and the ratio at which different mutation types can arise. 
Although still uncertain about how much of the genome is directly associated with the distinguished phenotypes, 
GWAS has indicated clusters of loci (linkage groups? Operon?) along the genome to be causative. 
To mimic this and compare dynamics of neutral vs. effect under selection, 
we create regions of effect. 
In these genomic elements of effect, we uniformly allow additive, recessive, and dominant, mutation that can push phenotype in either the positive or negative direction to arise. 
Every genomic position outside of these effect regions only allows for selectively neutral mutations to arise.

@@

@@

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/FST_HRR.png}
  		\caption{Distributions of $F_{st}$  values between the three subpopulations throughout the simulations.
		On the left, $F_{st}$ was calculated only on neutral mutation frequencies. On the right, $F_{st}$ was calculated
		only on effect (Impact on phenotype) mutation frequencies. 
		Effect mutations are acted upon by selection as they have an affect on fitness where neutral mutations  }
  		\label{fig:Fst}
	\end{center}
\end{figure}

@@

%FIGURES THAT SHOW THIS:
%-Time to adaptation
\begin{figure}
	\begin{center}
  		\includegraphics[width=0.8\linewidth]{matplotlibPlots/TimeToAdaptation.png}
  		\caption{Time to adaptation as a function of migration rate ($M$) parameter value. This is where we measure how many generation
		it takes for the introduced population's mean phenotype to come within 0.5 of the original lakes average phenotype. 
		Each point represents a simulation run at some value of $M$. 
		The yellow dashed line is the average of all points at each respective parameter value}
  		\label{fig:TimeToAdaptation}
	\end{center}
\end{figure}

@@

%FIGURES THAT SHOW THIS:
%-MPAA / IND
%-Total/Avg shared
\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MPFAI.png}
  		\caption{Distributions of mean percentage of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
		freshwater adapted alleles.
		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
		while the percentage tells us the variance of the haplotype}
		\label{fig:MPFAI}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MFAI.png}
  		\caption{Distributions of mean number of freshwater adapted alleles (FAA) per individual throughout the simulation run, for each subpopulation.
		We count the total number of freshwater alleles for each individual before averaging them in each population and dividing by the total number of defined
		freshwater adapted alleles.
		Looking at total number of FAA per individual gives us an idea behind how many alleles underly a freshwater haplotype, 
		}
  		\label{fig:MPFAI}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{plotlyPlots/NumFAA.png}
  		\caption{(PLACEHOLDER)}
		\label{fig:NumFAA}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{plotlyPlots/StandingGeneticVariation.png}
  		\caption{(PLACEHOLDER)}
		\label{fig:NumFAA}
	\end{center}
\end{figure}

@@

%FIGURES THAT SHOW THIS:
%- Phenotype Distribution
\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{matplotlibPlots/MeanPhenotype1.png}
  		\caption{Distribution of mean phenotype throughout simulation runs at separate migration rate ($M$) parameter values, for each population.
		The dashed pink line (Pheno = +10) is the optimum phenotype any individual in the marine environment.
		In contrast the purple line at (Pheno = -10) represents the optimum for any individual in the freshwater environment. 
		All individuals at generation 0 (beginning of the simulation) 
		}
  		\label{fig:MeanPhenotype}
	\end{center}
\end{figure}

@@

\subsection{Genomic Architecture}

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{matplotlibPlots/Power_FP.png}
  		\caption{ 
		Statistical Power and False Positives as a function of $F_{st}$ threshold. 
		Statistical Power is the likelihood that a SNP will be predicted to have an effect on phenotype when there is an effect to be detected (?).
		False Positives give us the ratio of SNPs that effect phenotype to total SNPs greater than the $F_{st}$ threshold.
		}
  		\label{fig:Power_FP}
	\end{center}
\end{figure}

Here, we take a closer look at the genomic architecture of local adaptation between freshwater and marine individuals. 
Often when looking at real datas, biologists observe patterns of $F_{st}$ and clustering of alleles that suggest causative loci for certain traits.
However, researchers would like to know how parameters such as migration and recombination impact the results from natural populations.
To give an idea about the impact of these parameters in our model, we examine the distribution, location, and false positive rate on predicting effect loci
given the $F_{st}$ per SNP across the genome.

\subsubsection*{False-Positives \& Statistical Power}

Given that migration increases the gene flow between subpopulations, how valid are $F_{st}$ peaks at different $M$. 
Knowing exactly which mutations effect phenotype in our simulations, 
we can look at the statistical power and false positives given $F_{st}$ per SNP across the genome. 
In Figure \ref{fig:Power_FP} , looking at an $F_{st}$ threshold greater than 1, we see the two lowest migration rates $10^{-5}$ and $10^{-4}$ having very little statical power. 
This along with low false positive rate across all $F_{st}$ threshold values is fairly predictable when you consider the high $F_{st}$ values across the genome.

%The lowest migration rate of $10^{-5}$ also has a very low false positive rate across all $F_{st}$ threshold values.

%\subsubsection*{Low Recombination Causes Clustering}
%\begin{comment}
\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-5.png}
  		\caption{ (PLACEHOLDER)
		}
  		\label{fig:Fst1}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-4.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Fst2}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-3.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Fst3}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/FstAcross5e-2.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Fst4}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-5.png}
  		\caption{ (PLACEHOLDER)
		}
  		\label{fig:Haplo1}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-4.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Haplo2}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-3.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Haplo3}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/Haplo5e-2.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Haplo4}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-5.png}
  		\caption{ (PLACEHOLDER)
		}
  		\label{fig:Haplo1}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-4.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Haplo2}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-3.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Haplo3}
	\end{center}
\end{figure}

\begin{figure}[h!tb]
	\begin{center}
  		\includegraphics[width=0.7\linewidth]{plotlyPlots/PhenotypeThroughout5e-2.png}
  		\caption{(PLACEHOLDER)
		}
  		\label{fig:Haplo4}
	\end{center}
\end{figure}
%\end{comment}
